---
title: "Activity Patterns"
subtitle: "Project 1 in Reproducible Research Course"
author: "Nicolas Moreno"
date: "May 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 1, digits = 2)
```

## Loading and preprocessing the data

The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

```{r load, cache=FALSE, warning=FALSE, message=FALSE}
setwd("~/Desktop/DataScience/RepData_PeerAssessment1/")
if(!file.exists("activity.csv")){
system("unzip activity.zip")
}
activity <- read.csv("activity.csv")
library(dplyr)
library(ggplot2)
```

## Mean Total Number of Steps per Day

```{r meantotalsteps}

totalsteps <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
totalstepsmean <- round(mean(totalsteps$steps),2)
totalstepsmedian <- median(totalsteps$steps)

```

```{r totalstepshistogram}
ggplot(data=totalsteps, aes(steps)) +
  geom_histogram(breaks=seq(0,25000,by=5000)) 
```

The mean total number of steps is `r round(totalstepsmean,2)` and the median is `r totalstepsmedian`.

## Average Daily Activity Pattern 

```{r computetime, echo=FALSE}
time <- format(Sys.time(), "%a %b %d %X %Y")
rand <- rnorm(1)
```

The current time is `r time`. My favorite random number is `r rand`.
```{r simulate data, echo=TRUE}
x <- rnorm(100); y <- x + rnorm(100, sd = 0.5)
```

```{r scatterplot, fig.height=4}
par(mar= c(5,4,1,1), las =1)
plot(x,y, main = "My simulated Data")
```

